---
title       : Chapter 6, The Capital-Labor Split in the Twenty-First Century
subtitle    : (Dynamics of the Capital/Income Ratio)
author      : Patrick Toche
framework   : html5slides  # {minimal, io2012, html5slides, shower, dzslides, ...}
hitheme     : solarized_light  # {tomorrow, solarized_light, zenburn...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3, libraries/morris, libraries/highcharts]}
github      : {user: patricktoche, repo: piketty}
url         : {lib: "../../libraries", assets: "../../assets"}
css         : "../../assets/css/custom.css" # tested
lecnum      : "06"

---

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

### Introduction  

- Chapter 5 analyzes the capital/income ratio.  

- Chapter 6 analyzes the division of national income between labor and capital.  

- The first fundamental law of capitalism:  
$$
\alpha = r \times \beta
$$
gives a correspondence between the capital/income ratio and the division of income between labor and capital.  

---

### The First Fundamental Law of Capitalism

- The first fundamental law of capitalism gives a correspondence between the capital/income ratio and the division of income between labor and capital.  
- For example, if the capital stock is equal to six years of national income ($\beta$ = 6), and if the average return on capital is 5 percent a year (r = 5%), then the share of income from capital $\alpha$ (as a percentage of national income) is 30 percent, and the share of income from labor is therefore 70 percent.  

- The central question is: How is the rate of return on capital determined?  


--- 

### Evolution of the Capital Income Share

- The capital income share $\alpha$ has followed a U-shaped curve similar to that of the capital/income ratio $\beta$, although the depth of the U is less pronounced.  
- Changes in the rate of return on capital $r$ have mitigated the evolution of the quantity of capital $\beta$:  
    - $r$ is higher in periods when $\beta$ is lower.  
    - $r$ is lower in periods when $\beta$ is higher.  
- The capital shares and rates of return are calculated by adding the various amounts of income from capital included in national accounts, including rents, profits, dividends, interest, royalties, etc., but excluding interest on public debt and excluding taxes.  

---

### Capital Income Share and Capital/Income Ratio

- The capital income share $\alpha$ was about 35–40 percent in both Britain and France in the late 18th and throughout the 19th century.  
- In the middle of the 20th century $\alpha$ fell to 20–25 percent, then rose again to 25–30 percent in the late 20th and early 21st.  
- This corresponds to an average rate of return on capital $r$ of about 5–6 percent in the 18th and 19th centuries,  7–8 percent in the mid-20th century, 4–5 percent in the late 20th and early 21st.  

---

### Capital Income Share and Capital/Income Ratio

- The yield on the riskiest assets (including industrial capital in the form of partnerships in family firms and shares of stock in listed corporations) is often greater than 7–8 percent.  
- The yield on less risky assets is about 4–5 percent for farmland in the 18th and 19th centuries and 3–4 percent for real estate in the early 21st century. 
- These are pre-tax returns.

---

### The Return on Capital in Historical Perspective
  

- There has been no pronounced long-term trend either upward or downward.  
- The pure return rose significantly above 6 percent following the massive destruction of property and numerous shocks to capital in the two world wars, but subsequently returned fairly rapidly to the lower levels observed in the past.   

---

### The capital-labor split in Britain, 1770-2010  
<figure>  
<img src = "../../figures/Figure_6_1.png", alt = "Figure 6.1. The capital-labor split in Britain, 1770-2010">  
<figcaption class = 'figcaption'>Figure 6.1. During the nineteenth century, capital income (rent, profits, dividends, interest...) absorbed about 40 percent of national income versus 60 percent for labor income (including both wage and non-wage income).</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The capital-labor split in Britain, 1770-2010  

```{r 'Figure_6_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_1.Rda") 
# round data for rChart tooltip display
df_6_1$value <- round(df_6_1$value, 2)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_1, type = 'lineChart') 
n$yAxis(axisLabel = 'Labor and capital income (% national income)')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$chart(shapes = c('circle', 'square'))
n
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_1.html', standalone = TRUE)
```
<iframe src = "figures/Figure_6_1.html" alt = "Figure 6.1. The capital-labor split in Britain, 1770-2010">
<figcaption class = 'figcaption'>Figure 6.1. During the nineteenth century, capital income (rent, profits, dividends, interest...) absorbed about 40 percent of national income versus 60 percent for labor income (including both wage and non-wage income).</figcaption>
</iframe>

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---  

### The capital-labor split in France, 1820-2010  

<figure>  
<img src = "../../figures/Figure_6_2.png", alt = "Figure 6.2. The capital-labor split in France, 1820-2010">  
<figcaption class = 'figcaption'>Figure 6.2. In the twenty-first century, capital income (rent, profits, dividends, interest...) absorbs about 30 percent of national income versus 70 percent for labor income (including both wage and non-wage income).</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The capital-labor split in France, 1820-2010  

```{r 'Figure_6_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_2.Rda") 
# round data for rChart tooltip display
df_6_2$value <- round(df_6_2$value, 2)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_2, type = 'lineChart') 
n$yAxis(axisLabel = 'Labor and capital income (% national income)')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_2.html', alt = "Figure 6.2. The capital-labor split in France, 1820-2010">
<figcaption class = 'figcaption'>Figure 6.2. In the twenty-first century, capital income (rent, profits, dividends, interest...) absorbs about 30 percent of national income versus 70 percent for labor income (including both wage and non-wage income).</figcaption>
</iframe>  

<footer class = 'footnote'>  
Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The pure rate of return on capital in Britain, 1770-2010  

<figure>  
<img src = "../../figures/Figure_6_3.png", alt = "Figure 6.3. The pure rate of return on capital in Britain, 1770-2010">  
<figcaption class = 'figcaption'>Figure 6.3. The pure rate of return to capital is roughly stable around 4-5 percent in the long run.</figcaption>  
</figure>  

<footer class = 'footnote'>
Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The pure rate of return on capital in Britain, 1770-2010 

```{r 'Figure_6_3_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_3.Rda") 
# round data for rChart tooltip display
df_6_3$value <- round(df_6_3$value, 3)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_3, type = 'lineChart') 
n$yAxis(axisLabel = 'Annual rate of return')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_3.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_3.html', alt = "Figure 6.3. The pure rate of return on capital in Britain, 1770-2010">
<figcaption class = 'figcaption'>Figure 6.3. The pure rate of return to capital is roughly stable around 4-5 percent in the long run.</figcaption>
</iframe>  

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The pure rate of return on capital in France, 1820-2010  

<figure>  
<img src = "../../figures/Figure_6_4.png", alt = "Figure 6.4. The pure rate of return on capital in France, 1820-2010.">  
<figcaption class = 'figcaption'>Figure 6.4. The observed average rate of return displays larger fluctuations than the pure rate of return during the twentieth century.</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The pure rate of return on capital in France, 1820-2010

```{r 'Figure_6_4_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_4.Rda") 
# round data for rChart tooltip display
df_6_4$value <- round(df_6_4$value, 3)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_4, type = 'lineChart') 
n$yAxis(axisLabel = 'Annual rate of return')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_4.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_4.html', alt = "The pure rate of return on capital in France, 1820-2010.">
<figcaption class = 'figcaption'>Figure 6.4. The observed average rate of return displays larger fluctuations than the pure rate of return during the twentieth century.</figcaption>
</iframe>  

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The capital share in rich countries, 1975-2010  

<figure>  
<img src = "../../figures/Figure_6_5.png", alt = "Figure 6.5. The capital share in rich countries, 1975-2010">  
<figcaption class = 'figcaption'>Figure 6.5. Capital income absorbs between 15 percent and 25 percent of national income in rich countries in 1970, and between 25 percent and 30 percent in 2000-2010.</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The capital share in rich countries, 1975-2010  

```{r 'Figure_6_5_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_5.Rda") 
# round data for rChart tooltip display
df_6_5$value <- round(df_6_5$value, 2)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_5, type = 'lineChart') 
n$yAxis(axisLabel = 'Capital income (% national income)')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_5.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_5.html', alt = "Figure 6.5. The capital share in rich countries, 1975-2010">
<figcaption class = 'figcaption'>Figure 6.5. Capital income absorbs between 15 percent and 25 percent of national income in rich countries in 1970, and between 25 percent and 30 percent in 2000-2010.</figcaption>
</iframe>  

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The profit share in the value added of corporations in France, 1900–2010  

<figure>  
<img src = "../../figures/Figure_6_6.png", alt = "Figure 6.6. The profit share in the value added of corporations in France, 1900–2010">  
<figcaption class = 'figcaption'>Figure 6.6. The share of gross profits in gross value added of corporations rose from 25 percent in 1982 to 33 percent in 2010; the share of net profits in net value added rose from 12 percent to 20 percent.</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The profit share in the value added of corporations in France, 1900–2010  

```{r 'Figure_6_6_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_6.Rda") 
# round data for rChart tooltip display
df_6_6$value <- round(df_6_6$value, 2)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_6, type = 'lineChart') 
n$yAxis(axisLabel = 'Profit share in value added')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_6.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_6.html', alt = "Figure 6.6. The profit share in the value added of corporations in France, 1900–2010">
<figcaption class = 'figcaption'>Figure 6.6. The share of gross profits in gross value added of corporations rose from 25 percent in 1982 to 33 percent in 2010; the share of net profits in net value added rose from 12 percent to 20 percent.</figcaption>
</iframe>  

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The share of housing rent in national income in France, 1900-2010   

<figure>  
<img src = "../../figures/Figure_6_7.png", alt = "Figure 6.7. The share of housing rent in national income in France, 1900-2010.">  
<figcaption class = 'figcaption'>Figure 6.7. The share of housing rent (rental value of dwellings) rose from 2 percent of national income in 1948 to 10 percent in 2010.</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The share of housing rent in national income in France, 1900-2010   

```{r 'Figure_6_7_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_7.Rda") 
# round data for rChart tooltip display
df_6_7$value <- round(df_6_7$value, 3)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_7, type = 'lineChart') 
n$yAxis(axisLabel = 'Share of housing rent in national income')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_7.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_7.html', alt = "Figure 6.7. The share of housing rent in national income in France, 1900-2010.">
<figcaption class = 'figcaption'>Figure 6.7. The share of housing rent (rental value of dwellings) rose from 2 percent of national income in 1948 to 10 percent in 2010.</figcaption>
</iframe>  

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The capital share in national income in France, 1900-2010    

<figure>  
<img src = "../../figures/Figure_6_8.png", alt = "Figure 6.8. The capital share in national income in France, 1900-2010.">  
<figcaption class = 'figcaption'>Figure 6.8. The share of capital income (net profits and rents) rose from 15 percent of national income in 1982 to 27 percent in 2010.</figcaption>  
</figure>  

<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The capital share in national income in France, 1900-2010   

```{r 'Figure_6_8_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_6_8.Rda") 
# round data for rChart tooltip display
df_6_8$value <- round(df_6_8$value, 3)
n <- nPlot(value ~ Year, group = 'variable', data = df_6_8, type = 'lineChart') 
n$yAxis(axisLabel = 'Profit share in value added')
n$chart(margin = list(left = 100)) # margin makes room for label
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline=TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_6_8.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_6_8.html', alt = "The capital share in national income in France, 1900-2010.">
<figcaption class = 'figcaption'>Figure 6.8. The share of capital income (net profits and rents) rose from 15 percent of national income in 1982 to 27 percent in 2010.</figcaption>
</iframe>  

<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---
