---
title       : Chapter 12, Global Inequality of Wealth in the Twenty-First Century
subtitle    : Thomas Piketty, Capital in the 21st Century (Harvard University Press 2014)
author      : Patrick Toche
framework   : io2012  # {minimal, io2012, html5slides, shower, dzslides, ...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3]}  # libraries/morris, libraries/highcharts, libraries/rickshaw, libraries/dimple
github      : {user: patricktoche, repo: piketty}
url         : {lib: "../../libraries", assets: "../../assets"}
lecnum      : "12"
--- 

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

### UNDER CONSTRUCTION

- These slides are under construction.
- Some of the figures are completed.
- Some of the figures require further work to fix bugs.
- Please check back later.

---

### Introduction  

- Chapter 11 analyzed the relative roles of inheritance and saving in capital formation in the long term.

- Chapter 12 studies the dynamics of wealth inequality at the global level.

---

### The world's billionaires according to Forbes
<figure class = "captioned">  
<img src = "../../figures/Figure_12_1.png" alt = "Figure 12.1. The world's billionaires according to Forbes, 1987-2013.">  
<figcaption class = 'figcaption'>Figure 12.1. Between 1987 and 2013, according to Forbes, the number of $ billionaires rose from 140 to 1,400, and their total wealth rose from 300 to 5,400 billion dollars.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The world's billionaires according to Forbes

```{r 'Figure_12_1_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}
# Script for formatting multiChart's y-axes scale and tick mark format
dualAxisChartScript <- 
  "<script>
      $(document).ready(function(){
        draw{{chartId}}()
      });

      function draw{{chartId}}(){  
        var opts = {{{ opts }}},
        data = {{{ data }}}

        if (!(opts.type==='pieChart')) {
          var data = d3.nest()
          .key(function(d){
            return opts.group === undefined ? 'main' : d[opts.group]
          })
          .entries(data);
        }

        //loop through to give an expected x and y
        //then give the type and yAxis hopefully provided by R
        data.forEach(
          function(variables) {
            variables.values.forEach(
              function(values){
                values.x = values[opts.x];
                values.y = values[opts.y];
              }
            );
            variables.type = opts.multi[variables.key].type;
            variables.yAxis = opts.multi[variables.key].yAxis;
          }
        );

        nv.addGraph(function() {
          var chart = nv.models[opts.type]()
          //.x(function(d) { return d[opts.x] })
          //.y(function(d) { return d[opts.y] })
          // Shrink the width a bit such that both axes' tick marks can fit
          .width(opts.width - 20)
          .height(opts.height)

          {{{ chart }}}

          {{{ xAxis }}}

          {{{ x2Axis }}}

          // force the y Axis range or limits for both axes
          chart.yDomain1([0, 6000]);
          chart.yDomain2([0, 1500]);

          // format ticks so 10000 appears as 10,000
          // prepend a currency symbol next to the formatted value
          var format = d3.format('0,0.0f');
          chart.yAxis1.tickFormat(function(y) {return '$' + format(y) + 'B';});
          
          // make space for axis labels
          chart.margin({top: 20, right: 100, bottom: 20, left: 100});

          // set axis labels
          chart.yAxis1.axisLabel('Wealth of $ billionaires in billions of dollars');
          chart.yAxis2.axisLabel('Number of $ billionaires (head count)');
          
          // show legend
          chart.showLegend(true);
          
          d3.select('#' + opts.id)
            .append('svg')
            .datum(data)
            .transition().duration(500)
            .call(chart);

          nv.utils.windowResize(chart.update);
          return chart;
        });
      };
  </script>"
require(rCharts)
load("../../data/df_12_1.Rda") 
# round data for rChart tooltip display
df_12_1$value <- round(df_12_1$value, 3)
# make multiChart to support dual axis
n <- nPlot(data = df_12_1, value ~ Year, group = 'variable', type = 'multiChart')
# Set which axes the item should follow
n$params$multi <- list(
  'Number of $ billionaires' = list(type = "line", yAxis = 2),
  'Total wealth in billions of dollars' = list(type = "line", yAxis = 1))
n$setTemplate(script = dualAxisChartScript)
n$xAxis(axisLabel = NULL)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_12_1.html', standalone = TRUE)
```
<iframe src = "figures/Figure_12_1.html" alt = "Figure 12.1. The world's billionaires according to Forbes, 1987-2013.">
</iframe><icaption class = 'icaption'>Figure 12.1. Between 1987 and 2013, according to Forbes, the number of $ billionaires rose from 140 to 1,400, and their total wealth rose from 300 to 5,400 billion dollars.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---  

### Billionaires as a fraction of global population and wealth

<figure class = "captioned">  
<img src = "../../figures/Figure_12_2.png" alt = "Figure 12.2. Billionaires as a fraction of global population and wealth, 1987-2013.">  
<figcaption class = 'figcaption'>Figure 12.2. Between 1987 and 2013, the number of billionaires per 100 million adults rose from five to thirty, and their share in aggregate private wealth rose from 0.4 percent to 1.5 percent.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Billionaires as a fraction of global population and wealth

```{r 'Figure_12_2_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE} 
# Script for formatting multiChart's y-axes scale and tick mark format
dualAxisChartScript <- 
  "<script>
      $(document).ready(function(){
        draw{{chartId}}()
      });

      function draw{{chartId}}(){  
        var opts = {{{ opts }}},
        data = {{{ data }}}

        if (!(opts.type==='pieChart')) {
          var data = d3.nest()
          .key(function(d){
            return opts.group === undefined ? 'main' : d[opts.group]
          })
          .entries(data);
        }

        //loop through to give an expected x and y
        //then give the type and yAxis hopefully provided by R
        data.forEach(
          function(variables) {
            variables.values.forEach(
              function(values){
                values.x = values[opts.x];
                values.y = values[opts.y];
              }
            );
            variables.type = opts.multi[variables.key].type;
            variables.yAxis = opts.multi[variables.key].yAxis;
          }
        );

        nv.addGraph(function() {
          var chart = nv.models[opts.type]()
          //.x(function(d) { return d[opts.x] })
          //.y(function(d) { return d[opts.y] })
          // Shrink the width a bit such that both axes' tick marks can fit
          .width(opts.width - 20)
          .height(opts.height)

          {{{ chart }}}

          {{{ xAxis }}}

          {{{ x2Axis }}}

          // force the y Axis range or limits for both axes
          chart.yDomain1([0, .016]);
          chart.yDomain2([0, 40]);

          // format ticks so 10000 appears as 10,000
          // prepend a currency symbol next to the formatted value
          var format = d3.format('0,0.0f');
          chart.yAxis1.tickFormat(function(y) {return Math.round(y*100*100)/100 + '%'});

          // make space for axis labels
          chart.margin({top: 20, right: 100, bottom: 20, left: 100});

          // set axis labels
          chart.yAxis1.axisLabel('Wealth of $ billionaires (% aggregate private wealth)');
          chart.yAxis2.axisLabel('Number of $ billionaires per 100 million adults');
          
          // show legend
          chart.showLegend(true);
          
          d3.select('#' + opts.id)
            .append('svg')
            .datum(data)
            .transition().duration(500)
            .call(chart);

          nv.utils.windowResize(chart.update);
          return chart;
        });
      };
  </script>"
require(rCharts)
load("../../data/df_12_2_rc.Rda") 
df_12_2 <- df_12_2_rc
# round data for rChart tooltip display
df_12_2$value <- round(df_12_2$value, 4)
# make multiChart to support dual axis
n <- nPlot(data = df_12_2, value ~ Year, group = 'variable', type = 'multiChart')
# Set which axes the item should follow
n$params$multi <- list(
  'Number of billionaires per 100 million adults' = list(type = "line", yAxis = 2),
  'Total wealth of billionaires, as a fraction of aggregate private wealth' = list(type = "line", yAxis = 1))
n$setTemplate(script = dualAxisChartScript)
n$xAxis(axisLabel = NULL)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_12_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_12_2.html' alt = "Figure 12.2. Billionaires as a fraction of global population and wealth, 1987-2013.">
</iframe><icaption class = 'icaption'>Figure 12.2. Between 1987 and 2013, the number of billionaires per 100 million adults rose from five to thirty, and their share in aggregate private wealth rose from 0.4 percent to 1.5 percent.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The share of top wealth fractiles in world wealth

<figure class = "captioned">  
<img src = "../../figures/Figure_12_3.png" alt = "Figure 12.3. The share of top wealth fractiles in world wealth, 1987-2013.">  
<figcaption class = 'figcaption'>Figure 12.3. Between 1987 and 2013, the share of the top 1/20 million fractile rose from 0.3 percent to 0.9 percent of world wealth, and the share of the top 1/100 million fractile rose from 0.1 percent to 0.4 percent.</figcaption>  
</figure> 
<footer class = 'footnote'>
Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The share of top wealth fractiles in world wealth

```{r 'Figure_12_3_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_12_3.Rda") 
# round data for rChart tooltip display
df_12_3$value <- round(df_12_3$value, 4)
n <- nPlot(data = df_12_3, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, .01))
n$chart(margin = list(left = 80))
n$yAxis(axisLabel = 'Annual value of bequests and gifts (% national income)')
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_12_3.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_12_3.html' alt = "Figure 12.3. The share of top wealth fractiles in world wealth, 1987-2013.">
</iframe><icaption class = 'icaption'>Figure 12.3. Between 1987 and 2013, the share of the top 1/20 million fractile rose from 0.3 percent to 0.9 percent of world wealth, and the share of the top 1/100 million fractile rose from 0.1 percent to 0.4 percent.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The world capital/income ratio

<figure class = "captioned">  
<img src = "../../figures/Figure_12_4.png" alt = "Figure 12.4. The world capital/income ratio, 1870-2100.">  
<figcaption class = 'figcaption'>Figure 12.4. According to the simulations (central scenario), the world capital/income ratio might be near to 700 percent by the end of the twenty-first century.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The world capital/income ratio

```{r 'Figure_12_4_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_12_4.Rda") 
# round data for rChart tooltip display
df_12_4$value <- round(df_12_4$value, 2)
n <- nPlot(data = df_12_4, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(2, 7))
n$chart(margin = list(left = 80))
n$yAxis(axisLabel = 'Value of private capital (% world income)')
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_12_4.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_12_4.html' alt = "Figure 12.4. The world capital/income ratio, 1870-2100.">
</iframe><icaption class = 'icaption'>Figure 12.4. According to the simulations (central scenario), the world capital/income ratio might be near to 700 percent by the end of the twenty-first century.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The distribution of world capital

<figure class = "captioned">  
<img src = "../../figures/Figure_12_5.png" alt = "Figure 12.5. The distribution of world capital, 1870-2100.">  
<figcaption class = 'figcaption'>Figure 12.5. According to the central scenario, Asian countries should own about half of world capital by the end of the twenty-first century.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The distribution of world capital

```{r 'Figure_12_5_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_12_5.Rda")
# round data for rChart tooltip display
df_12_5$value <- round(df_12_5$value, 3)
n <- nPlot(data = df_12_5, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1870, 2100), clipEdge = TRUE)
n$chart(yDomain = c(0, 26))  # use yDomain instead of forceY
n$chart(margin = list(left = 80)) 
n$yAxis(axisLabel = 'value of private capital (% world income)')
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 450, width = 900)
n$save('figures/Figure_12_5.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_12_5.html' alt = "Figure 12.5. The distribution of world capital, 1870-2100.">
</iframe><icaption class = 'icaption'>Figure 12.5. According to the central scenario, Asian countries should own about half of world capital by the end of the twenty-first century.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The net foreign asset position of rich countries

<figure class = "captioned">  
<img src = "../../figures/Figure_12_6.png" alt = "Figure 12.6. The net foreign asset position of rich countries.">  
<figcaption class = 'figcaption'>Figure 12.6. Unregistered financial assets held in tax havens are higher than the official net foreign debt of rich countries.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The net foreign asset position of rich countries

```{r 'Figure_12_6_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_12_6.Rda")
# round data for rChart tooltip display
df_12_6 <- df_12_6
df_12_6$value <- round(df_12_6$value, 3)
n <- nPlot(data = df_12_6, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(-.08, .1))
n$chart(margin = list(left = 80)) 
n$yAxis(axisLabel = 'Net foreign assets (% world output)')
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_12_6.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_12_6.html' alt = "Figure 12.6. The net foreign asset position of rich countries.">
</iframe><icaption class = 'icaption'>Figure 12.6. Unregistered financial assets held in tax havens are higher than the official net foreign debt of rich countries.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

