---
title       : Piketty's Capital
subtitle    : Selected figures and tables
author      : Patrick Toche
framework   : bootstrap  # {minimal, bootstrap, io2012, html5slides, shower, dzslides, ...}
hitheme     : solarized_light  # {tomorrow, solarized_light...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : []   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3, libraries/morris, libraries/highcharts]}
---
[https://github.com/ptoche/piketty](http://htmlpreview.github.io/?https://github.com/ptoche/piketty/blob/master/index.html)

### Notes on Thomas Piketty's Capital

```{r 'gsub', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(rCharts)
# set cache = FALSE for this chunk, so that it is run every time
knitr::knit_hooks$set(document = function(doc){
  gsub('`` ', '```', doc)
})
```

```{r 'setup', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
opts_chunk$set(echo = FALSE, cache = FALSE)
source('shared/shared.R')
```

- This is a website to display material related to Thomas Piketty, Capital in the 21st Century (Harvard University Press, 2014), in particular data, tables and figures. 
- This project was started in January 2015 and is expected to be updated on a weekly basis until May 2015.

---

## Content

```{r echo = FALSE, results = 'asis', cache = TRUE}
lectures <- dir('lectures', full = TRUE)
titles <- c(
  "Income and Output", 
  "Growth: Illusions and Realities", 
  "The Metamorphoses of Capital", 
  "From Old Europe to the New World",
  "The Capital/Income Ratio Over the Long Run",
  "The Capital-Labor Split in the Twenty-First Century",
  "Inequality and Concentration: Preliminary Bearings",
  "Two Worlds",
  "Inequality of Labor Income",
  "Inequality of Capital Ownership",
  "Merit and Inheritance in the Long Run",
  "Global Inequality of Wealth in the Twenty-First Century",
  "A Social State for the Twenty-First Century",
  "Rethinking the Progressive Income Tax", 
  "A Global Tax on Capital",
  "The Question of the Public Debt")
titles <- titles[1:6]
links <- paste0(seq_along(lectures), ". ", 
  "[", basename(lectures), " : ", titles, "]", "(", lectures, "/index.html)"
)
writeLines(links)
```

---

### Reproduce
This website is based on the slidify framework. The source files are hosted on github and you may download the whole project as a zip file or fork it to a github repository in your account. To compile this website from the `.Rmd` source, configure the `site.yml` and run `blogify(".")` from the root of the folder hierarchy. Slidify will automatically detect all the `Rmd` files, slidify them, and create all lecture slides along with the index page. 

You will need the `slidify`, `slidifyLibraries` and ``rCharts`, packages.

    install.packages("devtools")  
    devtools::install_github(c("slidify", "slidifyLibraries", "rCharts"), 
        "ramnathv", ref = "dev")  

In addition, I often use the following packages: ``ggplot2``, ``scales``, ``reshape2``, ``plyr``, so you may need to run this for each package:

    install.packages("ggplot2")

Write a short slide deck and publish it to [Github](http://www.github.com) with a few clicks using [RStudio](http://www.rstudio.com), with a single line of code.

<p><a class="btn" href="start.html">Learn More &raquo;</a></p>


### Layout

The `lectures` folder holds the lecture slides, one subfolder per lecture. The site's architecture looks like this:

```
assets      -> put custom img/js/css/layout assets
lectures    -> folder containing lectures
  Chapter 01
  Chapter 02
  Chapter 03
libraries   -> frameworks, highlighters and widgets
index.Rmd   -> Rmd source for home page
site.yml    -> Site related configuration
```

