---
title       : Chapter 8 -- Two Worlds
subtitle    : Thomas Piketty, Capital in the 21st Century (Harvard University Press 2014)
author      : Patrick Toche
framework   : io2012  # minimal, io2012, html5slides, shower, dzslides
mode        : selfcontained  # selfcontained, standalone, draft
widgets     : mathjax   # [mathjax, quiz, bootstrap]
ext_widgets : {rCharts: [libraries/nvd3]}  # libraries/morris, libraries/highcharts, libraries/rickshaw, libraries/dimple
url         : {lib: "../../libraries", assets: "../../assets"}
lecnum      : "08"
--- 

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

### UNDER CONSTRUCTION

- These slides are under construction.
- Some of the figures are completed.
- Some of the figures require further work to fix bugs.
- Please check back later.

---
### Introduction  

- Chapter 7 analyzes inequality and distribution at the individual level.

- Chapter 8 reviews the historical evolution of inequality around the world.  

---

### Income inequality in France
<figure class = "captioned">  
<img src = "../../figures/Figure_8_1.png" alt = "Figure 8.1. Income inequality in France, 1910-2010.">  
<figcaption class = 'figcaption'>Figure 8.1. Inequality of total income (labor and capital) has dropped in France during the twentieth century, while wage inequality has remained the same.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Income inequality in France

```{r 'Figure_8_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_1.Rda") 
# round data for rChart tooltip display
df_8_1$value <- round(df_8_1$value, 3)
n <- nPlot(data = df_8_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.2, .5))
n$yAxis(axisLabel = 'Share of top decile in total (incomes or wages)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_1.html', standalone = TRUE)
```
<iframe src = "figures/Figure_8_1.html" alt = "Figure 8.1. Income inequality in France, 1910-2010.">
</iframe><icaption class = 'icaption'>Figure 8.1. Inequality of total income (labor and capital) has dropped in France during the twentieth century, while wage inequality has remained the same.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---  

### The fall of rentiers in France

<figure class = "captioned">  
<img src = "../../figures/Figure_8_2.png" alt = "Figure 8.2. The fall of rentiers in France, 1910-2010.">  
<figcaption class = 'figcaption'>Figure 8.2. The fall in the top percentile share (the top 1 percent highest incomes) in France between 1914 and 1945 is due to the fall of top capital incomes.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The fall of rentiers in France

```{r 'Figure_8_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_2.Rda") 
# round data for rChart tooltip display
df_8_2$value <- round(df_8_2$value, 3)
n <- nPlot(data = df_8_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.04, .22))
n$yAxis(axisLabel = 'Share in total income of various fractiles')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_8_2.html' alt = "Figure 8.2. The fall of rentiers in France, 1910-2010.">
</iframe><icaption class = 'icaption'>Figure 8.2. The fall in the top percentile share (the top 1 percent highest incomes) in France between 1914 and 1945 is due to the fall of top capital incomes.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The composition of top incomes in France in 1932

<figure class = "captioned">  
<img src = "../../figures/Figure_8_3.png" alt = "Figure 8.3.
The composition of top incomes in France in 1932.">  
<figcaption class = 'figcaptioncompact'>Figure 8.3. Labor income becomes less and less important as one goes up within the top decile of total income. Notes: (i) P90-95 includes individuals between percentiles 90 to 95, P95-99 includes the next 4 percent, P99-99.5 the next 0.5 percent, etc.; (ii) Labor income: wages, bonuses, pensions. Capital income: dividends, interest, rent. Mixed income: self-employment income.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The composition of top incomes in France in 1932

```{r 'Figure_8_3_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_3.Rda") 
# round data for rChart tooltip display
df_8_3$value <- round(df_8_3$value, 3)
df_8_3$Levels <- as.numeric(df_8_3$Quantile)
n <- nPlot(data = df_8_3, value ~ Levels, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1))
n$yAxis(axisLabel = 'Share in total income of various fractiles')
n$chart(margin = list(left = 100, right = 40))
n$xAxis(tickFormat = "#! function (x) { 
        ticks = ['P90-95', 'P95-99', 'P99-99.5', 'P99.5-99.9', 'P99.9-99.99', 'P99.99-100']
        return ticks[x-1] ;
} !#")
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_3.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_8_3.html' alt = "Figure 8.3. The composition of top incomes in France in 1932.">
</iframe><icaption class = 'icaptioncompact'>Figure 8.3. Labor income becomes less and less important as one goes up within the top decile of total income. Notes: (i) P90-95 includes individuals between percentiles 90 to 95, P95-99 includes the next 4 percent, P99-99.5 the next 0.5 percent, etc.; (ii) Labor income: wages, bonuses, pensions. Capital income: dividends, interest, rent. Mixed income: self-employment income.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The composition of top incomes in France in 2005

<figure class = "captioned">  
<img src = "../../figures/Figure_8_4.png" alt = "Figure 8.4. The composition of top incomes in France in 2005.">  
<figcaption class = 'figcaption'>Figure 8.4. Capital income becomes dominant at the level of the top 0.1 percent in France in 2005, as opposed to the top 0.5 percent in 1932.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The composition of top incomes in France in 2005

```{r 'Figure_8_4_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
# highcharts alternative
# h <- hPlot(data = df_8_4, x = 'Quantile', y = 'value', group = 'variable', color = 'variable', shape = 'variable', type = 'line')
require(rCharts)
load("../../data/df_8_4.Rda") 
# round data for rChart tooltip display
df_8_4$value <- round(df_8_4$value, 3)
df_8_4$Levels <- as.numeric(df_8_4$Quantile)
n <- nPlot(data = df_8_4, value ~ Levels, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1))
n$yAxis(axisLabel = 'Share in total income of various fractiles')
n$chart(margin = list(left = 100, right = 40))
n$xAxis(tickFormat = "#! function (x) { 
        ticks = ['P90-95', 'P95-99', 'P99-99.5', 'P99.5-99.9', 'P99.9-99.99', 'P99.99-100']
        return ticks[x-1] ;
} !#")
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_4.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_8_4.html' alt = "Figure 8.4. The composition of top incomes in France in 2005.">
</iframe><icaption class = 'icaption'>Figure 8.4. Capital income becomes dominant at the level of the top 0.1 percent in France in 2005, as opposed to the top 0.5 percent in 1932.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Income inequality in the United States

<figure class = "captioned">  
<img src = "../../figures/Figure_8_5.png" alt = "Figure 8.5. Income inequality in the United States, 1910-2010.">  
<figcaption class = 'figcaption'>Figure 8.5. The top decile share in U.S. national income dropped from 45-50% in the 1910s-1920s to less than 35% in the 1950s (this is the
1950 1960 fall documented by Kuznets); it then rose from less than 35% in the 1970s to 45-50% in the 2000s-2010s.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Income inequality in the United States

```{r 'Figure_8_5_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_5.Rda") 
# round data for rChart tooltip display
df_8_5$value <- round(df_8_5$value, 3)
n <- nPlot(data = df_8_5, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.3, .5))
n$yAxis(axisLabel = 'Share of top decile in national income')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_5.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_8_5.html' alt = "Figure 8.5. Income inequality in the United States, 1910-2010.">
</iframe><icaption class = 'icaption'>Figure 8.5. The top decile share in U.S. national income dropped from 45-50% in the 1910s-1920s to less than 35% in the 1950s (this is the
1950 1960 fall documented by Kuznets); it then rose from less than 35% in the 1970s to 45-50% in the 2000s-2010s.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Decomposition of the top decile in the United States

<figure class = "captioned">  
<img src = "../../figures/Figure_8_6.png" alt = "Figure 8.6. Decomposition of the top decile, United States, 1910-2010..">  
<figcaption class = 'figcaption'>Figure 8.6. The rise of the top decile income share since the 1970s is mostly due to the top percentile.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Decomposition of the top decile in the United States

```{r 'Figure_8_6_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_6.Rda") 
# round data for rChart tooltip display
df_8_6$value <- round(df_8_6$value, 3)
n <- nPlot(data = df_8_6, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.08, .26))
n$yAxis(axisLabel = 'Share of the different groups in total income')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_6.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_8_6.html' alt = "Figure 8.6. Decomposition of the top decile, United States, 1910-2010.">
</iframe><icaption class = 'icaption'>Figure 8.6. The rise of the top decile income share since the 1970s is mostly due to the top percentile.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### High incomes and high wages in the United States

<figure class = "captioned">  
<img src = "../../figures/Figure_8_7.png" alt = "Figure 8.7. High incomes and high wages in the United States, 1910-2010.">  
<figcaption class = 'figcaption'>Figure 8.7. The rise of income inequality since the 1970s is largely due to the rise of wage inequality.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### High incomes and high wages in the United States

```{r 'Figure_8_7_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_7.Rda") 
# round data for rChart tooltip display
df_8_7$value <- round(df_8_7$value, 3)
n <- nPlot(data = df_8_7, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.2, .5))
n$yAxis(axisLabel = 'Share of top decile in total (incomes or wages)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_7.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_8_7.html' alt = "Figure 8.7. High incomes and high wages in the United States, 1910-2010.">
</iframe><icaption class = 'icaption'>Figure 8.7. The rise of income inequality since the 1970s is largely due to the rise of wage inequality.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The transformation of the top 1 percent in the United States

<figure class = "captioned">  
<img src = "../../figures/Figure_8_8.png" alt = "Figure 8.8. The transformation of the top 1 percent in the United States, 1910-2010.">  
<figcaption class = 'figcaption'>Figure 8.8. The rise in the top 1 percent highest incomes since the 1970s is largely due to the rise in the top 1 percent highest wages.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The transformation of the top 1 percent in the United States

```{r 'Figure_8_8_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_8.Rda") 
# round data for rChart tooltip display
df_8_8$value <- round(df_8_8$value, 3)
n <- nPlot(data = df_8_8, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.04, .26))
n$yAxis(axisLabel = 'Share of top decile in total (incomes or wages)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_8.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_8_8.html' alt = "Figure 8.8. The transformation of the top 1 percent in the United States, 1910-2010.">
</iframe><icaption class = 'icaption'>Figure 8.8. The rise in the top 1 percent highest incomes since the 1970s is largely due to the rise in the top 1 percent highest wages.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The composition of top incomes in the United States in 1929.

<figure class = "captioned">  
<img src = "../../figures/Figure_8_9.png" alt = "Figure 8.9. The composition of top incomes in the United States in 1929.">  
<figcaption class = 'figcaption'>Figure 8.9. Labor income becomes less and less important as one moves up within the top income decile.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The composition of top incomes in the United States in 1929.

```{r 'Figure_8_9_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_9.Rda") 
# round data for rChart tooltip display
df_8_9$value <- round(df_8_9$value, 3)
df_8_9$Levels <- as.numeric(df_8_9$Quantile)
n <- nPlot(data = df_8_9, value ~ Levels, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1))
n$yAxis(axisLabel = 'Share in total income of various fractiles')
n$chart(margin = list(left = 100, right = 40))
n$xAxis(tickFormat = "#! function (x) { 
        ticks = ['P90-95', 'P95-99', 'P99-99.5', 'P99.5-99.9', 'P99.9-99.99', 'P99.99-100']
        return ticks[x-1] ;
} !#")
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_9.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_8_9.html' alt = "Figure 8.9. The composition of top incomes in the United States in 1929.">
</iframe><icaption class = 'icaption'>Figure 8.9. Labor income becomes less and less important as one moves up within the top income decile.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The composition of top incomes in the United States in 2007.

<figure class = "captioned">  
<img src = "../../figures/Figure_8_10.png" alt = "Figure 8.10. The composition of top incomes in the United States in 2007.">  
<figcaption class = 'figcaption'>Figure 8.10. Capital income becomes dominant at the level of top 0.1 percent in 2007, as opposed to the top 1 percent in 1929.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The composition of top incomes in the United States in 2007.

```{r 'Figure_8_10_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_8_10.Rda") 
# round data for rChart tooltip display
df_8_10$value <- round(df_8_10$value, 3)
df_8_10$Levels <- as.numeric(df_8_10$Quantile)
n <- nPlot(data = df_8_10, value ~ Levels, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1))
n$yAxis(axisLabel = 'Share in total income of various fractiles')
n$chart(margin = list(left = 100, right = 40))
n$xAxis(tickFormat = "#! function (x) { 
        ticks = ['P90-95', 'P95-99', 'P99-99.5', 'P99.5-99.9', 'P99.9-99.99', 'P99.99-100']
        return ticks[x-1] ;
} !#")
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_8_10.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_8_10.html' alt = "Figure 8.10. The composition of top incomes in the United States in 2007.">
</iframe><icaption class = 'icaption'>Figure 8.10. Capital income becomes dominant at the level of top 0.1 percent in 2007, as opposed to the top 1 percent in 1929.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  


