---
title       : Introduction
author      : Patrick Toche
framework   : html5slides  # {minimal, io2012, html5slides, shower, dzslides, ...}
hitheme     : solarized_light  # {tomorrow, solarized_light, zenburn...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3, libraries/morris, libraries/highcharts]}
github      : {user: patricktoche, repo: piketty}
url         : {lib: "../../libraries", assets: "../../assets"}
css         : "../../assets/css/custom.css" # tested
lecnum      : "01"

--- {class: title-slide}

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

---

### Introduction

- This set of lecture notes surveys selected topics from *Capital in the Twenty-First Century*, a book written by economist Thomas Piketty, published in English in 2014 to great acclaim.
- The intended audience for these notes is any curious person with or without training in economics.
- The book deals with technical matters but its streamlined presentation of technical issues makes it highly readable.

---

### Thomas Piketty
<br />
<figure class = "centeredImage">  
<img src = "../../images/Piketty_BloombergBusinessWeek_2014.jpg" alt = "Thomas Thomas Piketty (born 1971)." style="width: auto; height: auto; max-height: 500px; ">
<figcaption class = 'figcaption'>Piketty's Capital took America by storm. 
</figcaption>  
</figure> 

---

### Capital in the 21st Century

- What do we know about the distribution of income? 
- What do we know about the distribution of wealth? 
- What prediction can we make about the evolution of income and wealth over the long term?
- Do the forces of private capital accumulation inevitably lead to the concentration of wealth in ever fewer hands?
- Can the balancing forces of growth, competition, and technological progress be strong enough to reduce inequality and reduce class conflicts?
- Chapter 1 lays out some basic facts about population and income, reviews the basic concepts of national income accounting, and introduces the basic forces behind the evolution of income and wealth over the long term.

--- 

### Thomas Robert Malthus
<br />
<figure class = "centeredImage">  
<img src = "../../images/Thomas_Robert_Malthus.jpg" alt = "Thomas Robert Malthus (1766-1834)." style="width: auto; height: auto; max-height: 500px;">
<figcaption class = 'figcaption'>Reverend Thomas Malthus (1766-1834).
</figcaption>  
</figure> 

---

### Malthus and the Problem of Overpopulation

- Social science research on the distribution of wealth was for a long time based on a relatively limited set of firmly established facts together with a wide variety of purely theoretical speculations. 
- In 1798, Thomas Malthus published his *Essay on the Principle of Population*. His main fear was overpopulation.
- Population in the French kingdom had increased from 20 million in 1700 to 30 million by 1780. This unprecedentedly rapid population growth contributed to a stagnation of agricultural wages and an increase in land rents in the decades prior to the Revolution of 1789.
- Malthus argued that all welfare assistance to the poor must be halted and that reproduction by the poor must be reduced to avoid overpopulation, chaos and misery.

---

### David Ricardo

<br />
<figure class = "centeredImage">  
<img src = "../../images/David_Ricardo_by_Thomas_Phillips.jpg" alt = "David Ricardo (1772-1823)." style="width: auto; height: auto; max-height: 500px;">
<figcaption class = 'figcaption'>David Ricardo (1772-1823), Portrait by Thomas Phillips.
</figcaption>  
</figure> 

---

### Ricardo and the Principle of Scarcity

- In 1817, Ricardo published his *Principles of Political Economy and Taxation*. His main concern was the long-term evolution of land prices and land rents. 
- Ricardo thought capitalism suffered from an internal contradiction. 
- As both population and output grow steadily, land tends to become increasingly scarce relative to other goods. The law of supply and demand then implies that the price of land will rise continuously, as will the rents paid to landlords. The landlords will therefore claim a growing share of national income, at the expense of the rest of the population.
- For Ricardo, a politically acceptable solution was to impose a steadily increasing tax on land rents.

---

### Ricardo and the Principle of Scarcity

- Ricardo's prediction proved wrong: in the course of economic development, as the share of agriculture in national income decreased, the value of farm land declined relative to other forms of wealth.
- Ricardo had not anticipated the importance of technological progress and its role in offsetting the principle of scarcity.
- The principle of scarcity remains important: replace the price of farmland with the price of urban real estate and consider what would happen to wealth inequality if the price trend of 1970–2010 were extrapolated to 2010–2050!
- The mechanism of supply and demand eventually counterbalances extreme trends: If the supply of some good is short and its price too high, then demand for that good would decrease and eventually its price would decrease.

---

### Karl Marx

<br />
<figure class = "centeredImage">  
<img src = "../../images/Karl_Marx_1875.jpg" alt = "Karl Marx (1818-1883)." style="width: auto; height: auto; max-height: 500px;">
<figcaption class = 'figcaption'>Karl Marx (1818-1883), Portrait dated 1875.
</figcaption>  
</figure> 

---

### Marx and the Kapital

- In 1867, Marx published the first volume of *Capital*. The other two volumes would be published posthumously by Friedrich Engels, the co-author of the *Communist Manifesto* (1848).
- The question was no longer whether farmers could feed a growing population or whether land prices would go on rising.
- During the first half of the 19th c. wages stagnated at very low levels, in some cases even lower than previous centuries. 
- The capital income share increased considerably in Britain and France in the first half of the 19th c.
- With the industrial revolution and the vast rural exodus workers crowded into urban slums. 
- Poverty was now visible in urban areas side by side with extreme opulence. 

---

### Marx and the Falling Rate of Profit

- Marx predicted a tendency for capitalists to accumulate wealth and become concentrated in ever fewer hands, with no natural limit to the process. 
- Marx predicted an excessive accumulation of capital would lead to a falling rate of profit and the collapse of industrial capitalism. 
- Marx's prediction did not come true. In the last third of the nineteenth century, wages finally began to increase.
- The communist revolution came in the most backward country in Europe, Russia, and not in the advanced industrial countries.

---

### Simon Kuznets and the Kuznets Curve

- In 1953, Kuznets published *Shares of Upper Income Groups in Income and Savings*, a study of the United States over 1913–1948. 
- Kuznets analyzed the evolution of the share of each decile and upper centiles of income. He noted a sharp reduction in income inequality in the U.S. between 1913 and 1948. 
- In 1913, the top 10 percent of US earners received 45–50 percent of annual national income. In 1948, 30–35 percent.
- The 'Kuznets curve': inequality increases in the early stages of industrialization, decreases later. 
- However, the compression of high incomes was largely accidental, caused by the Great Depression and World War II.

---

### Income inequality in the United States

<figure class = "centeredImage">  
<img src = "../../figures/Figure_0_1.png" alt = "Figure I.1. Income inequality in the United States, 1910-2012.">  
<figcaption class = 'figcaption'>Figure I.1. The top decile share in U.S. national income dropped from 45-50% in the 1910s-1920s to less than 35% in the 1950s (this is the 1950-1960 fall documented by Kuznets); it then rose from less than 35% in the 1970s to 45-50% in the 2000s-2010s.
</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Income inequality in the United States

```{r 'Figure_0_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_0_1.Rda") 
# round data for rChart tooltip display
df_0_1$value <- round(df_0_1$value, 3)
n <- nPlot(data = df_0_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.3, .5))
n$yAxis(axisLabel = 'Share of top decile in national income')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(showLegend = FALSE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_0_1.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_0_1.html' alt = "Figure I.1. Income inequality in the United States, 1910-2012.">
</iframe><icaption class = 'icaption'>Figure I.1. The top decile share in U.S. national income dropped from 45-50% in the 1910s-1920s to less than 35% in the 1950s (this is the 1950-1960 fall documented by Kuznets); it then rose from less than 35% in the 1970s to 45-50% in the 2000s-2010s.</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The capital-income ratio in Europe

<figure class = "centeredImage">  
<img src = "../../figures/Figure_0_2.png" alt = "Figure I.2. The growth of world population, 1700-2012.">  
<figcaption class = 'figcaption'>Figure I.2. Aggregate private wealth was worth about 6-7 years of national income in Europe in 1910, between 2 and 3 years in 1950, and between 4 and 6 years in 2010.</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The capital-income ratio in Europe

```{r 'Figure_0_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_0_2.Rda") 
# round data for rChart tooltip display
df_0_2$value <- round(df_0_2$value, 3)
n <- nPlot(data = df_0_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(1.5, 7.5))
n$yAxis(axisLabel = 'Market value of private capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  }
</style>')
n$save('figures/Figure_0_2.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_0_2.html' alt = "Figure I.2. The capital-income ratio in Europe, 1870-2012.">
</iframe><icaption class = 'icaption'>Figure I.2. Aggregate private wealth was worth about 6-7 years of national income in Europe in 1910, between 2 and 3 years in 1950, and between 4 and 6 years in 2010.</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

