---
title       : Chapter 3, The Metamorphoses of Capital 
subtitle    : Thomas Piketty, Capital in the 21st Century (Harvard University Press 2014)
author      : Patrick Toche
framework   : io2012  # {minimal, io2012, html5slides, shower, dzslides, ...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3]}  # libraries/morris, libraries/highcharts, libraries/rickshaw, libraries/dimple
github      : {user: patricktoche, repo: piketty}
url         : {lib: "../../libraries", assets: "../../assets"}
lecnum      : "03"
--- {class: cover-slide}

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

---

### Introduction  

- This set of slides surveys selected topics from *Capital in the Twenty-First Century*, a book written by economist Thomas Piketty, published in English in 2014 to great acclaim.
- All source files for this course are available for download by anyone without restrictions at https://github.com/ptoche/piketty
- The full course is expected to be completed by April 2015.
- Chapter 2 reviewed the main stages of income and output growth since the Industrial Revolution.
- Chapter 3 reviews the metamorphoses of capital in Britain and France since the eighteenth century.  

---

### The Transformation of Capital

- Figures 3.1 and 3.2 show the capital structure of the UK and France.
    - `1`. The capital/income ratio $K/Y$ followed an inverted-bell curve ('U-shape'). Stable in the 18th and 19th centuries at about 600% or 700%; very large decline to 200% or 300% in the 20th during the depression and WWII; going back to the pre-WWI level after WWII.
    - `2`. The composition of capital $K$ has transformed over the very long run: agricultural land has been replaced by buildings, business capital, and financial capital invested in firms and government organizations.
- The depression, the wars, and inflation during 1914-1945 wiped the slate clean and created the illusion of virtuous, merit-based capitalism &mdash; the 'myth of the Kuznets curve' derided by Piketty.

---

### The Structure of Capital

- **3 categories of capital**: `1` farm land; `2` residential land & housing; `3` other types including land and buildings used for business, infrastructure, machinery, computers, patents. 

    - **National capital $=$ farmland $+$ housing $+$ other domestic capital $+$ net foreign capital**  
- In 1700, the total value of farmland represented 400% to 500% of national income, nearly 2/3 of total national capital and 3/4 of all economic activity and employment. The value of housing was 100% of national income; the 'other' category, the same as housing.
- In 2010, farmland is less than 10% of national income, less than 2% of capital. Housing is 300% of national income; like the 'other' category.

---

### The Transformation of Capital

There are 2 main causes behind this structural transformation:

- `1`. growing importance of housing quality and value (driven by urbanization and agglomeration for production purposes); 

- `2`. growing importance of buildings and structures in the production of non-agricultural goods and services (machinery, warehouses, offices).

---

### The Rise and Fall of Foreign Capital

- The role of foreign capital in the economies of the United Kingdom and France follows the ups and downs of their colonial history. 
- Foreign possessions first became important around 1750–1800.
- Net assets owned in the rest of the world peaked at high levels in 1914, collapsed during 1914–1945, have since stabilized at a much lower level.
- In 1914, the UK had the largest colonial empire and owned foreign assets worth 200% of national income &mdash; 6 times the value of domestic farmland. Foreign assets were yielding 5% a year in dividends, interest, and rent. National income was 10% higher than domestic product. 
- France had the 2nd largest colonial empire, owned foreign assets worth more than 100%. National income was 6% higher than domestic product.

---

### The Rise of Foreign Capital

- These very large net foreign assets allowed the UK and France to run structural trade deficits. 
- During 1880-1914, both countries' trade deficits averaged 1–2% of national income. That is, both countries imported more goods and services from the rest of the world than they exported. 
- Because their income from foreign assets exceeded 5% of national income, their balance of payments was strongly positive and their holdings of foreign assets increased year after year!
- Thus, while the colonial powers consumed more than they produced, the colonized countries increased their debt to the colonies!

--- 

### The Fall of Foreign Capital

- Following 2 world wars, some local wars, 1 depression, high inflation, the British and French empires lost their colonies and foreign wealth.
- After 1950, their net foreign asset holdings have been close to zero.
- In 2010, net foreign assets are negligible, as they were in 1700. The ratio of total capital to national income is similar in both periods. 
- The most important long-run structural change is the fall of farmland and the rise of real estate and working capital. 

---

### Income and Capital Today

- The UK and France are typical rich countries.
- National income per capita is about 30,000 euros per year.
- National capital 180,000 euros per capita &mdash; 6 times national income. 
- Farmland is a few thousand euros per capita.
- National capital includes about 90,000 euros in housing and 90,000 euros of other domestic capital, such as financial instruments.

---

### Public & Private Capital

- Figures 3.3 and 3.4 shows public assets and public debt in the UK and France, and Figures 3.5 and 3.6 show public and private capital. 
- Public capital is the difference between assets and liabilities of the state.
- Private capital is the difference between assets and liabilities of people.
- National capital is the sum of public capital and private capital. 
- Public assets include non-financial wealth (public buildings such as schools, universities, hospitals) and financial wealth (stocks of private corporations, foreign currencies, gold).
- Public assets (financial and non-financial) are about 100% of national income in the UK and 150% in France. 
- Public debt are about 100% of national income in both countries.
- Net public wealth is close to zero in the UK, about 30% in France.

---

### Net Public Wealth

- If the French and British governments decided to sell off all their assets and pay off their debts, there would be 'nothing' left.
- But these estimates are imprecise. It is not easy to set a precise market value on schools and hospitals, on railway lines, on highways. 
- The difference between financial and non-financial assets can be arbitrary.
- When the French government privatized 'France Telecom' and the Post Office, the buildings and structures previously treated as non-financial assets became classified as financial assets.
- Private wealth accounts for 95-99% of national wealth in both countries. These are 'capitalist' &mdash; not communist &mdash; economies.

---

### Public Debt in British History

- The UK has a long history of borrowing 'to the hilt.'
- During 1700–1720, public debt reached 50% of national income. During 1760–1770, 100%. After the Napoleonic wars and after WWII, UK public debt exceeded 200% of GDP.
- It took a century of budget surpluses to gradually reduce public debt below 30% by 1914 &mdash; successive government did not repudiate it.
- Public debt lost its significance because national income outgrew it, mostly because of strong economic growth.

--- 

### Public Debt in British History

- The large rise in public debt of 1770–1810 was financed largely by a rise in private saving, so that national capital remained stable at around 700% of national income. 
- The deficits raised demand for private wealth bid up its return, making government bonds attractive relative to other assets. 
- From 1815 to 1914 &mdash; a century! &mdash; the UK government paid 5% of GDP every year to their (wealthy, willing and happy) creditors, financing this transfer of wealth with 'primary' surpluses (the primary surplus is tax revenue minus non-interest government spending).

---

### Public Debt in French History

- Following the 1789 Revolution and 'terreur', France defaulted on 2/3 of public debt ('banqueroute des deux tiers'), resorted to high inflation (episode of the 'assignats'), and by 1815 the public debt accumulated by the fallen aristocratic 'Ancien Régime' was below 20%.
- Revolutions are often followed by a default on public debt: the confederate states in 1865 (end of the American civil war), the Russian 'Soviets' in 1917 and Chinese 'Maoists' in 1949  (communist revolutions).

---

### Who's Afraid of the Public Debt?

- Europe, America, Japan are heavily in debt. A government bond is a claim of one segment of the population (those who receive interest) on another (those who pay taxes). Why should debt matter?
- In the 19th century, Marx observed rising interest rates and saw public debt as a 'lever' of the accumulation of private capital.
- Inflation in 1913–1950 averaged 13% a year (prices multiplied by 100) in France; 3% a year (prices multiplied by 3) in the UK. In 1950-1980 inflation averaged 6% (prices multiplied by 6) in the UK, 10% in the 1970s alone. Inflation shrunk the real value of debt and 'solved' the debt problem.
- Friedman saw public debt as a lever of public spending and wealth redistribution. So did Ronald Reagan and Margaret Thatcher.
- Marx saw a tool of the rich; Friedman saw a tool of the poor.
- Reagan was not afraid of debt. Nor was G.W.Bush.

---

### 'Ricardian Equivalence'

- In 1817, David Ricardo suggested that public debt does not always 'crowd out' private capital. 
- While British public debt was 200% of GDP, private investment was healthy &mdash; the public debt was financed by a rise in private saving.
- The gigantic public debt had no effect on national wealth.
- Public debt was no more than a claim of one group of people on another.
- Writing in 1936, John Maynard Keynes felt that the 'euthanasia of the rentier' would be the simplest way to reduce the burden of public debt.

--- 

### Who's Afraid of Inflation?

- Inflation is a powerful tool for wealth redistribution. 
- It destroys the value of bonds more than other forms of wealth. 
- It destroys the purchasing power of cash for rich and poor alike.
- It works only by surprising bond holders, who then learn to distrust government and demand higher nominal interest rates to compensate for inflation and its risk. 
- Once built into expectations inflation becomes persistent and costly.
- Eliminating inflation requires banking reforms and a strong recession.
- The Germans are afraid of inflation. So are the French.

---

### Who's Afraid of Government?

- The history of public assets is less erratic than that of public debt.
- The value of public assets rose from 50% of national income in the 17th century to 150% in France and 100% in the UK today. 
- This reflects the expansion of the state, in particular in the areas of health, education, transportation, and communication.
- Periods of privatization and nationalization have alternated. Reagan and Thatcher initiated a wave of privatization in the 1980s. In the UK: British Petroleum, British Aerospace, British Gas, British Steel, British Telecom, Sealink ferries, Rolls-Royce, British Rail. France, Japan, and of course the former Soviet Union also privatized on a grand scale.


<!--- 

### Capital in Britain

<figure class = "captioned">
<img src = "../../figures/Figure_3_1.png" alt = "Figure 3.1. Capital in Britain, 1700-2010.">  
<figcaption class = 'figcaption'>Figure 3.1. National capital is worth about seven years of national income in Britain in 1700 (including four in agricultural land).</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Capital in Britain

```{r 'Figure_3_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_3_1.Rda") 
# To do: dynamic range selection
# round data for rChart tooltip display
df_3_1$value <- round(df_3_1$value, 3)
n <- nPlot(data = df_3_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1700, 2012), clipEdge = TRUE)
n$chart(yDomain = c(0, 8))  # use yDomain instead of forceY
n$yAxis(axisLabel = 'Value of national capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 450, width = 900)
n$save('figures/Figure_3_1.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_3_1.html' alt = "Figure 3.1. Capital in Britain, 1700-2010.">
</iframe><icaption class = 'icaption'>Figure 3.1. National capital is worth about seven years of national income in Britain in 1700 (including four in agricultural land).</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

<!---

### Capital in France

<figure class = "captioned">
<img src = "../../figures/Figure_3_2.png" alt = "Figure 3.2. Capital in France, 1700-2010.">  
<figcaption class = 'figcaption'>Figure 3.2. National capital is worth almost seven years of national income in France in 1910 (including one invested abroad).</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Capital in France

```{r 'Figure_3_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_3_2.Rda") 
# To do: dynamic range selection
# round data for rChart tooltip display
df_3_2$value <- round(df_3_2$value, 3)
n <- nPlot(data = df_3_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1700, 2012), clipEdge = TRUE)
n$chart(yDomain = c(0, 8))  # use yDomain instead of forceY
n$yAxis(axisLabel = 'Value of national capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 450, width = 900)
n$save('figures/Figure_3_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_3_2.html' alt = "Figure 3.2. Capital in France, 1700-2010.">
</iframe><icaption class = 'icaption'>Figure 3.2. National capital is worth almost seven years of national income in France in 1910 (including one invested abroad).</icaption>
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

<!---

### Public wealth in Britain

<figure class = "captioned">
<img src = "../../figures/Figure_3_3.png" alt = "Figure 3.3. Public wealth in Britain, 1700-2010.">  
<figcaption class = 'figcaption'>Figure 3.3. Public debt surpassed two years of national income in 1950 (versus one year for public assets).</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Public wealth in Britain

```{r 'Figure_3_3_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_3_3.Rda") 
# round data for rChart tooltip display
df_3_3$value <- round(df_3_3$value, 3)
n <- nPlot(data = df_3_3, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 2.4))
n$yAxis(axisLabel = 'Public assets and debt (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_3_3.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_3_3.html' alt = "Figure 3.3. Public wealth in Britain, 1700-2010.">
</iframe><icaption class = 'icaption'>Figure 3.3. Public debt surpassed two years of national income in 1950 (versus one year for public assets).</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

<!---

### Public wealth in France

<figure class = "captioned">
<img src = "../../figures/Figure_3_4.png" alt = "Figure 3.4. Public wealth in France, 1700-2010.">  
<figcaption class = 'figcaption'>Figure 3.4. Public debt is about one year of national income in France in 1780 as well as in 1880 and in 2000-2010.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Public wealth in France

```{r 'Figure_3_4_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_3_4.Rda") 
# round data for rChart tooltip display
df_3_4$value <- round(df_3_4$value, 3)
n <- nPlot(data = df_3_4, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 2.4))
n$yAxis(axisLabel = 'Public assets and debt (% national income)')
n$chart(margin = list(left = 80))
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_3_4.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_3_4.html' alt = "Figure 3.4. Public wealth in France, 1700-2010.">
</iframe><icaption class = 'icaption'>Figure 3.4. Public debt is about one year of national income in France in 1780 as well as in 1880 and in 2000-2010.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

<!---

### Private and public capital in Britain

<figure class = "captioned">
<img src = "../../figures/Figure_3_5.png" alt = "Figure 3.5. Private and public capital in Britain, 1700-2010.">  
<figcaption class = 'figcaption'>Figure 3.5. In 1810, private capital is worth eight years of national income in Britain (versus seven years for national capital).</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Private and public capital in Britain

```{r 'Figure_3_5_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_3_5.Rda") 
# round data for rChart tooltip display
df_3_5$value <- round(df_3_5$value, 3)
n <- nPlot(data = df_3_5, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(-2, 9))
n$yAxis(axisLabel = 'National, private and public capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_3_5.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_3_5.html' alt = "Figure 3.5. Private and public capital in Britain, 1700-2010.">
</iframe><icaption class = 'icaption'>Figure 3.5. In 1810, private capital is worth eight years of national income in Britain (versus seven years for national capital).</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

<!---

### Private and public capital in France

<figure class = "captioned">
<img src = "../../figures/Figure_3_6.png" alt = "Figure 3.6. Private and public capital in France, 1700-2010.">  
<figcaption class = 'figcaption'>Figure 3.6. In 1950, public capital is worth almost one year of national income versus two years for private capital.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

-->

---

### Private and public capital in France

```{r 'Figure_3_6_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_3_6.Rda") 
# round data for rChart tooltip display
df_3_6$value <- round(df_3_6$value, 3)
n <- nPlot(data = df_3_6, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(-2, 9))
n$yAxis(axisLabel = 'National, private and public capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_3_6.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_3_6.html' alt = "Figure 3.6. Private and public capital in France, 1700-2010.">
</iframe><icaption class = 'icaption'>Figure 3.6. In 1950, public capital is worth almost one year of national income versus two years for private capital.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

