---
title       : Chapter 11, Merit and Inheritance in the Long Run
subtitle    : Thomas Piketty, Capital in the 21st Century (Harvard University Press 2014)
author      : Patrick Toche
framework   : io2012  # {minimal, io2012, html5slides, shower, dzslides, ...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3]}  # libraries/morris, libraries/highcharts, libraries/rickshaw, libraries/dimple
github      : {user: patricktoche, repo: piketty}
url         : {lib: "../../libraries", assets: "../../assets"}
lecnum      : "11"
--- 

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

### UNDER CONSTRUCTION

- These slides are under construction.
- Some of the figures are completed.
- Some of the figures require further work to fix bugs.
- Please check back later.

---

### Introduction  

- Chapter 10 reviewed the historical evolution of wealth inequality and the factors behind the historic compression of inequality.

- Chapter 11 analyzes the relative roles of inheritance and saving in capital formation in the long term.

---

<figure class = "centered">  
<img src = "../../tables/Table_11_1.png" alt = "The age-wealth profile in France, 1872-2010: Average wealth of each group (% of average wealth of 50 to 59 year-olds)" style="max-width: 1000px!important; max-height: 600px!important; margin: 0em; padding: 0em;">
</figure> 


---

### The annual inheritance flow as a fraction of national income
<figure class = "captioned">  
<img src = "../../figures/Figure_11_1.png" alt = "Figure 11.1. The annual inheritance flow as a fraction of national income, France, 1820-2010.">  
<figcaption class = 'figcaption'>Figure 11.1. The annual inheritance flow was about 20–25 percent of national income during the nineteenth century and until 1914; it then fell to less than 5 percent in the 1950s, and returned to about 15 percent in 2010.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The annual inheritance flow as a fraction of national income

```{r 'Figure_11_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_1.Rda") 
# round data for rChart tooltip display
df_11_1$value <- round(df_11_1$value, 3)
n <- nPlot(data = df_11_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, .26))
n$yAxis(axisLabel = 'Annual value of inheritance and gifts (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_1.html', standalone = TRUE)
```
<iframe src = "figures/Figure_11_1.html" alt = "Figure 11.1. The annual inheritance flow as a fraction of national income, France, 1820-2010.">
</iframe><icaption class = 'icaption'>Figure 11.1. The annual inheritance flow was about 20–25 percent of national income during the nineteenth century and until 1914; it then fell to less than 5 percent in the 1950s, and returned to about 15 percent in 2010.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---  

### The mortality rate in France

<figure class = "captioned">  
<img src = "../../figures/Figure_11_2.png" alt = "Figure 11.2. The mortality rate in France, 1820-2100.">  
<figcaption class = 'figcaption'>Figure 11.2. The mortality rate fell in France during the twentieth century (rise of life expectancy), and should increase somewhat during the twenty-first century (baby-boom effect).</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The mortality rate in France

```{r 'Figure_11_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_2.Rda") 
# round data for rChart tooltip display
df_11_2$value <- round(df_11_2$value, 6)
n <- nPlot(data = df_11_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.010, .024))
n$yAxis(axisLabel = 'Adult mortality rate (%)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_11_2.html' alt = "Figure 11.2. The mortality rate in France, 1820-2100.">
</iframe><icaption class = 'icaption'>Figure 11.2. The mortality rate fell in France during the twentieth century (rise of life expectancy), and should increase somewhat during the twenty-first century (baby-boom effect).</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Average age of decedents and inheritors

<figure class = "captioned">  
<img src = "../../figures/Figure_11_3.png" alt = "Figure 11.3. Average age of decedents and inheritors: France, 1820-2100.">  
<figcaption class = 'figcaption'>Figure 11.3. The average of (adult) decedents rose from less than 60 years to almost 80 years during the twentieth century, and the average age at the time of inheritance rose from 30 years to 50 years.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Average age of decedents and inheritors

```{r 'Figure_11_3_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_3.Rda") 
# round data for rChart tooltip display
df_11_3$value <- round(df_11_3$value, 1)
n <- nPlot(data = df_11_3, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(20, 90))
n$yAxis(axisLabel = 'Average age in years')
n$chart(margin = list(left = 80)) 
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_3.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_11_3.html' alt = "Figure 11.3. Average age of decedents and inheritors: France, 1820-2100.">
</iframe><icaption class = 'icaption'>Figure 11.3. The average of (adult) decedents rose from less than 60 years to almost 80 years during the twentieth century, and the average age at the time of inheritance rose from 30 years to 50 years.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Inheritance flow versus mortality rate

<figure class = "captioned">  
<img src = "../../figures/Figure_11_4.png" alt = "Figure 11.4. Inheritance flow versus mortality rate: France, 1820-2010.">  
<figcaption class = 'figcaption'>Figure 11.4. The annual flow of inheritance (bequests and gifts) is equal to about 2.5 percent of aggregate wealth in 2000-2010 versus 1.2 percent for the mortality rate.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Inheritance flow versus mortality rate

```{r 'Figure_11_4_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_4.Rda") 
# round data for rChart tooltip display
df_11_4$value <- round(df_11_4$value, 3)
n <- nPlot(data = df_11_4, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.01, .04))
n$yAxis(axisLabel = 'Annual rate of transmission or mortality (%)')
n$chart(margin = list(left = 80))
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_4.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_11_4.html' alt = "Figure 11.4. Inheritance flow versus mortality rate: France, 1820-2010.">
</iframe><icaption class = 'icaption'>Figure 11.4. The annual flow of inheritance (bequests and gifts) is equal to about 2.5 percent of aggregate wealth in 2000-2010 versus 1.2 percent for the mortality rate.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Average wealth at death and average wealth of the living

<figure class = "captioned">  
<img src = "../../figures/Figure_11_5.png" alt = "Figure 11.5. The ratio between average wealth at death and average wealth of the living: France, 1820-2010.">  
<figcaption class = 'figcaption'>Figure 11.5. In 2000-2010, the average wealth at death is 20 percent higher than that of the living if one omits the gifts that were made before death, but more than twice as large if one re-integrates gifts.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Average wealth at death and average wealth of the living

```{r 'Figure_11_5_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
# To do: add legend title
legendtitle <- "Ratio of average wealth at death to average wealth of the decendents and the living"
require(rCharts)
load("../../data/df_11_5_rc.Rda") # df_11_5_rc has shorter names
# round data for rChart tooltip display
df_11_5 <- df_11_5_rc
df_11_5$value <- round(df_11_5$value, 3)
n <- nPlot(data = df_11_5, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.9, 2.3))
n$yAxis(axisLabel = 'Ratio between the average wealth of decedents and the living')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$addParams(title = legendtitle) # To do: how to add title
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_5.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_11_5.html' alt = "Figure 11.5. The ratio between average wealth at death and average wealth of the living: France, 1820-2010.">
</iframe><icaption class = 'icaption'>Figure 11.5. In 2000-2010, the average wealth at death is 20 percent higher than that of the living if one omits the gifts that were made before death, but more than twice as large if one re-integrates gifts.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Observed and simulated inheritance flow

<figure class = "captioned">  
<img src = "../../figures/Figure_11_6.png" alt = "Figure 11.6. Observed and simulated inheritance flow: France, 1820-2100.">  
<figcaption class = 'figcaption'>Figure 11.6. Simulations based upon the theoretical model indicate that the level of the inheritance flow in the twenty-first century will depend upon the growth rate and the net rate of return to capital.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Observed and simulated inheritance flow

```{r 'Figure_11_6_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_6_rc.Rda")  # df_11_6_rc has shorter names
# round data for rChart tooltip display
df_11_6 <- df_11_6_rc
df_11_6$value <- round(df_11_6$value, 3)
n <- nPlot(data = df_11_6, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.04, .26))
n$yAxis(axisLabel = 'Annual value of bequests and gifts (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_6.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_11_6.html' alt = "Figure 11.6. Observed and simulated inheritance flow: France, 1820-2100.">
</iframe><icaption class = 'icaption'>Figure 11.6. Simulations based upon the theoretical model indicate that the level of the inheritance flow in the twenty-first century will depend upon the growth rate and the net rate of return to capital.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The share of inherited wealth in total wealth

<figure class = "captioned">  
<img src = "../../figures/Figure_11_7.png" alt = "Figure 11.7. The share of inherited wealth in total wealth: France, 1850-2100.">  
<figcaption class = 'figcaption'>Figure 11.7. Inherited wealth represents 80-90 percent of total wealth in France in the nineteenth century; this share fell to 40-50 percent during the twentieth century, and might return to 80-90 percent during the twenty-first century.
</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The share of inherited wealth in total wealth

```{r 'Figure_11_7_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_7_rc.Rda") 
df_11_7 <- df_11_7_rc
# round data for rChart tooltip display
df_11_7$value <- round(df_11_7$value, 3)
n <- nPlot(data = df_11_7, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0.4, 1))
n$yAxis(axisLabel = 'Cumulated value of inherited wealth (% total wealth of the living)')
n$chart(margin = list(left = 80))
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_7.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_11_7.html' alt = "Figure 11.7. The share of inherited wealth in total wealth: France, 1850-2100.">
</iframe><icaption class = 'icaption'>Figure 11.7. Inherited wealth represents 80-90 percent of total wealth in France in the nineteenth century; this share fell to 40-50 percent during the twentieth century, and might return to 80-90 percent during the twenty-first century.
</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Annual inheritance flow and household disposable income

<figure class = "captioned">  
<img src = "../../figures/Figure_11_8.png" alt = "Figure 11.8. The annual inheritance flow as a fraction of household disposable income: France, 1820-2010.">  
<figcaption class = 'figcaption'>Figure 11.8. Expressed as a fraction of household disposable income (rather than national income), the annual inheritance flow is about 20 percent in 2010, in other words, close to its nineteenth-century level.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Annual inheritance flow and household disposable income

```{r 'Figure_11_8_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_8_rc.Rda") 
df_11_8 <- df_11_8_rc
# round data for rChart tooltip display
df_11_8$value <- round(df_11_8$value, 3)
n <- nPlot(data = df_11_8, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.02, .28))
n$yAxis(axisLabel = 'Annual value of inheritance and gifts (% household disposable income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_8.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_11_8.html' alt = "Figure 11.8. The annual inheritance flow as a fraction of household disposable income: France, 1820-2010.">
</iframe><icaption class = 'icaption'>Figure 11.8. Expressed as a fraction of household disposable income (rather than national income), the annual inheritance flow is about 20 percent in 2010, in other words, close to its nineteenth-century level.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The share of inheritance in the total resources

<figure class = "captioned">  
<img src = "../../figures/Figure_11_9.png" alt = "Figure 11.9. The share of inheritance in the total resources (inheritance and work) of cohorts born in 1790-2030.">  
<figcaption class = 'figcaption'>Figure 11.9. Inheritance made about 25 percent of the resources of nineteenth-century cohorts, down to less than 10 percent for cohorts born in 1910-1920 (who should have inherited in 1950-1960).</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The share of inheritance in the total resources

```{r 'Figure_11_9_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_9.Rda") 
# round data for rChart tooltip display
df_11_9$value <- round(df_11_9$value, 3)
n <- nPlot(data = df_11_9, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.08, .26))
n$yAxis(axisLabel = 'Share of inheritance of total resources of each cohort)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_9.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_11_9.html' alt = "Figure 11.9. The share of inheritance in the total resources (inheritance and work) of cohorts born in 1790-2030.">
</iframe><icaption class = 'icaption'>Figure 11.9. Inheritance made about 25 percent of the resources of nineteenth-century cohorts, down to less than 10 percent for cohorts born in 1910-1920 (who should have inherited in 1950-1960).</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The dilemma of Rastignac

<figure class = "captioned">  
<img src = "../../figures/Figure_11_10.png" alt = "Figure 11.10. The dilemma of Rastignac for cohorts born in 1790-2030.">  
<figcaption class = 'figcaption'>Figure 11.10. In the nineteenth century, the living standards that could be attained by the top 1 percent inheritors were a lot higher than those that could be attained by the top 1 percent labor earners.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The dilemma of Rastignac

```{r 'Figure_11_10_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_10_rc.Rda") 
df_11_10 <- df_11_10_rc
# round data for rChart tooltip display
df_11_10$value <- round(df_11_10$value, 3)
n <- nPlot(data = df_11_10, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(4, 30))
n$yAxis(axisLabel = 'Multiples of average income attained by bottom 50% of wage earners')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_10.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_11_10.html' alt = "Figure 11.10. The dilemma of Rastignac for cohorts born in 1790-2030.">
</iframe><icaption class = 'icaption'>Figure 11.10. In the nineteenth century, the living standards that could be attained by the top 1 percent inheritors were a lot higher than those that could be attained by the top 1 percent labor earners.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Inheritance in units of lifetime labor income

<figure class = "captioned">  
<img src = "../../figures/Figure_11_11.png" alt = "Figure 11.11. Which fraction of a cohort receives in inheritance the equivalent of a lifetime labor income?">  
<figcaption class = 'figcaption'>Figure 11.11. Within the cohorts born around 1970-1980, 12-14 percent of individuals receive in inheritance the equivalent of the lifetime labor income received by the bottom 50 percent less well paid workers.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Inheritance in units of lifetime labor income

```{r 'Figure_11_11_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_11.Rda") 
# round data for rChart tooltip display
df_11_11$value <- round(df_11_11$value, 3)
n <- nPlot(data = df_11_11, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.02, .16))
n$yAxis(axisLabel = 'Fraction of each cohort')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_11.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_11_11.html' alt = "Figure 11.11. Which fraction of a cohort receives in inheritance the equivalent of a lifetime labor income?">
</iframe><icaption class = 'icaption'>Figure 11.11. Within the cohorts born around 1970-1980, 12-14 percent of individuals receive in inheritance the equivalent of the lifetime labor income received by the bottom 50 percent less well paid workers.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### The inheritance flow in Europe

<figure class = "captioned">  
<img src = "../../figures/Figure_11_12.png" alt = "Figure 11.12. The inheritance flow in Europe, 1900-2010.">  
<figcaption class = 'figcaption'>Figure 11.12. The inheritance flow follows a U-shape in curve in France as well as in the United Kingdom and Germany. It is possible that gifts are underestimated in the United Kingdom at the end of the period.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### The inheritance flow in Europe

```{r 'Figure_11_12_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_11_12.Rda") 
# round data for rChart tooltip display
df_11_12$value <- round(df_11_12$value, 3)
n <- nPlot(data = df_11_12, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, .26))
n$yAxis(axisLabel = 'Annual value of bequests and gifts (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_11_12.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_11_12.html' alt = "Figure 11.12. The inheritance flow in Europe, 1900-2010.">
</iframe><icaption class = 'icaption'>Figure 11.12. The inheritance flow follows a U-shape in curve in France as well as in the United Kingdom and Germany. It is possible that gifts are underestimated in the United Kingdom at the end of the period.</icaption> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  
