---
title       : Chapter 13, A Social State for the Twenty-First Century; Chapter 14, Rethinking the Progressive Income Tax
subtitle    : Thomas Piketty, Capital in the 21st Century (Harvard University Press 2014)
author      : Patrick Toche
framework   : io2012  # {minimal, io2012, html5slides, shower, dzslides, ...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3]}  # libraries/morris, libraries/highcharts, libraries/rickshaw, libraries/dimple
url         : {lib: "../../libraries", assets: "../../assets"}
lecnum      : "11"
--- 

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

### UNDER CONSTRUCTION

- These slides are under construction.
- Some of the figures are completed.
- Some of the figures require further work to fix bugs.
- Please check back later.

---

### Introduction  

- Chapter 12 studied the dynamics of wealth inequality at the global level.

- Chapter 13 considers political institutions that might regulate today's global patrimonial capitalism fairly and efficiently.

- Chapter 14 examines more closely the structure of taxes and other government revenues.

---

### Tax revenues in rich countries

<figure class = "captioned">  
<img src = "../../figures/Figure_13_1.png" alt = "Figure 13.1. Tax revenues in rich countries, 1870-2010.">  
<figcaption class = 'figcaption'>Figure 13.1. Total tax revenues were less than 10 percent of national income in rich countries until 1900-1910; they represent between 30 percent and 55 percent of national income in 2000-2010.</figcaption>  
</figure> 
<footer class = 'footnote'>Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Tax revenues in rich countries

```{r 'Figure_13_1_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_13_1.Rda") 
# round data for rChart tooltip display
df_13_1$value <- round(df_13_1$value, 4)
n <- nPlot(data = df_13_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, .55))
n$chart(margin = list(left = 80))
n$yAxis(axisLabel = 'Tax revenue (% national income)')
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_13_1.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_13_1.html' alt = "Figure 13.1. Tax revenues in rich countries, 1870-2010.">
</iframe><icaption class = 'icaption'>Figure 13.1. Total tax revenues were less than 10 percent of national income in rich countries until 1900-1910; they represent between 30 percent and 55 percent of national income in 2000-2010.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Top income tax rates

<figure class = "captioned">  
<img src = "../../figures/Figure_14_1.png" alt = "Figure 14.1. Top income tax rates, 1900-2013.">  
<figcaption class = 'figcaption'>Figure 14.1. The top marginal tax rate of the income tax (applying to the highest incomes) in the U.S. dropped from 70% in 1980 to 28% in 1988.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Top income tax rates

```{r 'Figure_14_1_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_14_1.Rda") 
# round data for rChart tooltip display
df_14_1$value <- round(df_14_1$value, 2)
n <- nPlot(data = df_14_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1))
n$chart(margin = list(left = 80))
n$yAxis(axisLabel = 'Marginal tax rate applying to the highest incomes')
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_14_1.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_14_1.html' alt = "Figure 14.1. Top income tax rates, 1900-2013.">
</iframe><icaption class = 'icaption'>Figure 14.1. The top marginal tax rate of the income tax (applying to the highest incomes) in the U.S. dropped from 70% in 1980 to 28% in 1988.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Top inheritance tax rates

<figure class = "captioned">  
<img src = "../../figures/Figure_14_2.png" alt = "Figure 14.2. Top inheritance tax rates, 1900-2013.">  
<figcaption class = 'figcaption'>Figure 14.2. The top marginal tax rate of the inheritance tax (applying to the highest inheritances) in the U.S. dropped from 70% in 1980 to 35% in 2013.</figcaption>  
</figure> 
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Top inheritance tax rates

```{r 'Figure_14_2_rcharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_14_2.Rda")
# round data for rChart tooltip display
df_14_2$value <- round(df_14_2$value, 2)
n <- nPlot(data = df_14_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1))
n$chart(margin = list(left = 80))
n$yAxis(axisLabel = 'Marginal tax rate applying to the highest inheritances')
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = NULL)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 450, width = 900)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_14_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_14_2.html' alt = "Figure 14.2. Top inheritance tax rates, 1900-2013.">
</iframe><icaption class = 'icaption'>Figure 14.2. The top marginal tax rate of the inheritance tax (applying to the highest inheritances) in the U.S. dropped from 70% in 1980 to 35% in 2013.</icaption>
<footer class = 'footnote'>Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

