---
title       : Chapter 4, From Old Europe to the New World 
subtitle    : (The Dynamics of the Capital//Income Ratio)
author      : Patrick Toche
framework   : html5slides  # {minimal, io2012, html5slides, shower, dzslides, ...}
hitheme     : solarized_light  # {tomorrow, solarized_light, zenburn...}
mode        : selfcontained  # {selfcontained, standalone, draft}
widgets     : mathjax   # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3, libraries/morris, libraries/highcharts]}
github      : {user: patricktoche, repo: piketty}
url         : {lib: "../../libraries", assets: "../../assets"}
css         : "../../assets/css/custom.css" # tested
lecnum      : "04"

---

```{r 'preamble', message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, comment = NA, cache = FALSE}
require(knitr)
library(png)   # to display external png and control dimensions
library(grid)  # to display external png and control dimensions
opts_chunk$set(echo = FALSE, cache = FALSE)
source('../../shared/shared.R')
```

### Introduction  

- Chapter 3 reviews the metamorphoses of capital in Britain and France since the eighteenth century.  

- Chapter 4 extends this review to Germany, the United States and Canada, and addresses the question of slavery.

---

### Capital in Germany

<figure>  
<img src = "../../figures/Figure_4_1.png" alt = "Figure 4.1. Capital in Germany, 1870-2010.">  
<figcaption class = 'figcaption'>Figure 4.1. National capital is worth 6.5 years of national income in Germany in 1910 (including about 0.5 year invested abroad).</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Capital in Germany

```{r 'Figure_4_1_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_1.Rda") 
# To do: dynamic range selection
# round data for rChart tooltip display
df_4_1$value <- round(df_4_1$value, 3)
n <- nPlot(data = df_4_1, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1870, 2010), clipEdge = TRUE)
n$chart(yDomain = c(0, 8))  # use yDomain instead of forceY
n$yAxis(axisLabel = 'Value of national capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 500, width = 800)
n$save('figures/Figure_4_1.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_4_1.html' alt = "Figure 4.1. Capital in Germany, 1870-2010.">
</iframe><icaption class = 'icaption'>Figure 4.1. National capital is worth 6.5 years of national income in Germany in 1910 (including about 0.5 year invested abroad).</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Public wealth in Germany

<figure>  
<img src = "../../figures/Figure_4_2.png" alt = "Figure 4.2. Public wealth in Germany, 1870-2010.">  
<figcaption class = 'figcaption'>Figure 4.2. Public debt is worth almost one year of national income in Germany in 2010 (as much as assets).</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Public wealth in Germany

```{r 'Figure_4_2_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_2.Rda") 
# round data for rChart tooltip display
df_4_2$value <- round(df_4_2$value, 3)
n <- nPlot(data = df_4_2, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(.1, 1.1))
n$yAxis(axisLabel = 'Public assets and debt (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_4_2.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_4_2.html' alt = "Figure 4.2. Public wealth in Germany, 1870-2010.">
</iframe><icaption class = 'icaption'>Figure 4.2. Public debt is worth almost one year of national income in Germany in 2010 (as much as assets).</icaption>
<footer class = 'footnote'>  
Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Private and public capital in Germany

<figure>  
<img src = "../../figures/Figure_4_3.png" alt = "Figure 4.3. Private and public capital in Germany, 1870-2010.">  
<figcaption class = 'figcaption'>Figure 4.3. In 1970, public capital is worth almost one year of national income, versus slightly more than two for private capital.</figcaption>  
</figure> 
<footer class = 'footnote'>
Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Private and public capital in Germany

```{r 'Figure_4_3_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_3.Rda") 
# round data for rChart tooltip display
df_4_3$value <- round(df_4_3$value, 3)
n <- nPlot(data = df_4_3, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 7))
n$yAxis(axisLabel = 'National, private and public capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_4_3.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_4_3.html' alt = "Figure 4.3. Private and public capital in Germany, 1870-2010.">
</iframe><icaption class = 'icaption'>Figure 4.3. In 1970, public capital is worth almost one year of national income, versus slightly more than two for private capital.</icaption>
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Private and public capital in Europe

<figure>  
<img src = "../../figures/Figure_4_4.png" alt = "Figure 4.4. Private and public capital in Europe, 1870-2010.">  
<figcaption class = 'figcaption'>Figure 4.4. The fluctuations of national capital in Europe in the long run are mostly due to the fluctuations of private capital.</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Private and public capital in Europe

```{r 'Figure_4_4_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_4.Rda") 
# round data for rChart tooltip display
df_4_4$value <- round(df_4_4$value, 3)
n <- nPlot(data = df_4_4, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(-2, 8))
n$yAxis(axisLabel = 'Public and private capital (% national income)')
n$chart(margin = list(left = 80))
n$yAxis(tickFormat = "#! function(y) {return Math.round(y*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_4_4.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_4_4.html' alt = "Figure 4.4. Private and public capital in Europe, 1870-2010.">
</iframe><icaption class = 'icaption'>Figure 4.4. The fluctuations of national capital in Europe in the long run are mostly due to the fluctuations of private capital.</icaption>
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### National capital in Europe

<figure>  
<img src = "../../figures/Figure_4_5.png" alt = "Figure 4.5. National capital in Europe, 1870-2010.">  
<figcaption class = 'figcaption'>Figure 4.5. National capital (sum of public and private capital) is worth between two and three years of national income in Europe in 1950.</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### National capital in Europe

```{r 'Figure_4_5_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_5.Rda") 
# round data for rChart tooltip display
df_4_5$value <- round(df_4_5$value, 3)
n <- nPlot(data = df_4_5, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(1, 8))
n$yAxis(axisLabel = 'National capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_4_5.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_4_5.html' alt = "Figure 4.5. National capital in Europe, 1870-2010.">
</iframe><icaption class = 'icaption'>Figure 4.5. National capital (sum of public and private capital) is worth between two and three years of national income in Europe in 1950.</icaption>
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Capital in the United States

<figure>  
<img src = "../../figures/Figure_4_6.png" alt = "Figure 4.6. Capital in the United States, 1770-2010.">  
<figcaption class = 'figcaption'>Figure 4.6. National capital is worth three years of national income in the United States in 1770 (including 1.5 years in agricultural land).</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Capital in the United States

```{r 'Figure_4_6_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_6.Rda") 
# To do: dynamic range selection
# round data for rChart tooltip display
df_4_6$value <- round(df_4_6$value, 3)
n <- nPlot(data = df_4_6, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1770, 2012), clipEdge = TRUE)
n$chart(yDomain = c(0, 8))  # use yDomain instead of forceY
n$yAxis(axisLabel = 'Value of national capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 500, width = 800)
n$save('figures/Figure_4_6.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_4_6.html' alt = "Figure 4.6. Capital in the United States, 1770-2010.">
</iframe><icaption class = 'icaption'>Figure 4.6. National capital is worth three years of national income in the United States in 1770 (including 1.5 years in agricultural land).</icaption>
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Public wealth in the United States

<figure>  
<img src = "../../figures/Figure_4_7.png" alt = "Figure 4.7. Public wealth in the United States, 1770-2010.">  
<figcaption class = 'figcaption'>Figure 4.7. Public debt is worth one year of national income in the United States in 1950 (almost as much as assets).</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Public wealth in the United States

```{r 'Figure_4_7_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_7.Rda") 
# round data for rChart tooltip display
df_4_7$value <- round(df_4_7$value, 3)
n <- nPlot(data = df_4_7, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 1.3))
n$yAxis(axisLabel = 'Public assets and public debt (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_4_7.html', standalone = TRUE)
```
<iframe src = 'figures/Figure_4_7.html' alt = "Figure 4.7. Public wealth in the United States, 1770-2010.">
</iframe><icaption class = 'icaption'>Figure 4.7. Public debt is worth one year of national income in the United States in 1950 (almost as much as assets).</icaption>
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Private and public capital in the United States

<figure>  
<img src = "../../figures/Figure_4_8.png" alt = "Figure 4.8. Private and public capital in the United States, 1770-2010.">  
<figcaption class = 'figcaption'>Figure 4.8. In 2010, public capital is worth 20 percent of national income, versus over 400 percent for private capital.</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Private and public capital in the United States

```{r 'Figure_4_8_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_8.Rda") 
# round data for rChart tooltip display
df_4_8$value <- round(df_4_8$value, 3)
n <- nPlot(data = df_4_8, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'lineChart') 
n$chart(forceY = c(0, 5.5))
n$yAxis(axisLabel = 'National, private, and public capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$setTemplate(afterScript = '<style>
  .nv-point {
    stroke-opacity: 1!important;
    stroke-width: 5px!important;
    fill-opacity: 1!important;
  } 
</style>')
n$save('figures/Figure_4_8.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_4_8.html' alt = "Figure 4.8. Private and public capital in the United States, 1770-2010.">
</iframe><icaption class = 'icaption'>Figure 4.8. In 2010, public capital is worth 20 percent of national income, versus over 400 percent for private capital.</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Capital in Canada

<figure>  
<img src = "../../figures/Figure_4_9.png" alt = "Figure 4.9. Capital in Canada, 1860-2010.">  
<figcaption class = 'figcaption'>Figure 4.9. In Canada, a substantial part of domestic capital has always been held by the rest of the world, so that national capital has always been less than domestic capital.</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Capital in Canada

```{r 'Figure_4_9_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_9.Rda") 
# To do: dynamic range selection
# round data for rChart tooltip display
df_4_9$value <- round(df_4_9$value, 3)
n <- nPlot(data = df_4_9, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1860, 2010), clipEdge = TRUE)
n$chart(yDomain = c(0, 8))  # use yDomain instead of forceY
n$yAxis(axisLabel = 'Value of national capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 500, width = 800)
n$save('figures/Figure_4_9.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_4_9.html' alt = "Figure 4.9. Capital in Canada, 1860-2010.">
</iframe><icaption class = 'icaption'>Figure 4.9. In Canada, a substantial part of domestic capital has always been held by the rest of the world, so that national capital has always been less than domestic capital.</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Capital and slavery in the United States

<figure>  
<img src = "../../figures/Figure_4_10.png" alt = "Figure 4.10. Capital and slavery in the United States.">  
<figcaption class = 'figcaption'>Figure 4.10. The market value of slaves was about 1.5 years of US national income around 1770 (as much as land).</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Capital and slavery in the United States

```{r 'Figure_4_10_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_10.Rda") 
# To do: dynamic range selection
# round data for rChart tooltip display
df_4_10$value <- round(df_4_10$value, 3)
n <- nPlot(data = df_4_10, value ~ Year, group = 'variable', color = 'variable', shape = 'variable', type = 'stackedAreaChart') 
n$chart(xDomain = c(1770, 2010), clipEdge = TRUE)
n$chart(yDomain = c(0, 8))  # use yDomain instead of forceY
n$yAxis(axisLabel = 'Value of national capital (% national income)')
n$chart(margin = list(left = 80)) 
n$yAxis(tickFormat = "#! function(d) {return Math.round(d*100*100)/100 + '%'} !#")
n$xAxis(axisLabel = 'Year')
n$chart(style = 'stack')
n$chart(showControls = FALSE)
n$chart(useInteractiveGuideline = TRUE)
n$chart(color = c('#EFF3FF', '#BDD7E7', '#6BAED6', '#2171B5'))  # To do: automate color selection
n$addParams(height = 500, width = 800)
n$save('figures/Figure_4_10.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_4_10.html' alt = "Figure 4.10. Capital and slavery in the United States.">
</iframe><icaption class = 'icaption'>Figure 4.10. The market value of slaves was about 1.5 years of US national income around 1770 (as much as land).</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  

---

### Capital around 1770-1810: Old and New World.

<figure>  
<img src = "../../figures/Figure_4_11.png" alt = "Figure 4.11. Capital around 1770-1810: Old and New World.">  
<figcaption class = 'figcaption'>Figure 4.11. The combined value of agricultural land and slaves in the Southern United States surpassed four years of national income around 1770-1810.</figcaption>  
</figure> 
<footer class = 'footnote'>
Concept and data: Thomas Piketty. Chart created with ``ggplot2`` (author: Hadley Wickham)
</footer>  

---

### Capital around 1770-1810: Old and New World.

```{r 'Figure_4_11_rCharts',  message = FALSE, warning = FALSE, error = FALSE, echo = FALSE, tidy = FALSE, cache = FALSE}  
require(rCharts)
load("../../data/df_4_11.Rda") 
# round data for rChart tooltip display
df_4_11$value <- round(df_4_11$value, 3)
n <- nPlot(data = df_4_11, value ~ Country, group = 'variable', type = 'multiBarChart') 
n$yAxis(axisLabel = 'Value of capital (% national income)')
n$chart(margin = list(left = 80))
n$chart(color = colorPalette)
n$addParams(height = 500, width = 800)
n$save('figures/Figure_4_11.html', standalone = TRUE)
``` 
<iframe src = 'figures/Figure_4_11.html' alt = "Figure 4.11. Capital around 1770-1810: Old and New World.">
</iframe><icaption class = 'icaption'>Figure 4.11. The combined value of agricultural land and slaves in the Southern United States surpassed four years of national income around 1770-1810.</icaption> 
<footer class = 'footnote'>  
Concept and data: Thomas Piketty. Chart created with ``rCharts`` (author: Ramnath Vaidyanathan)  
</footer>  


